SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-07 back.tex

Kim {\em et. al.} proposed Signature Path Prefetcher (SPP)~\cite{SPP},
a confidence-based lookahead prefetcher.  SPP creates a signature
associated with a page address by compressing the history of
accesses. By correlating the signature with future likely delta
patterns, SPP learns both simple and complicated memory access
patterns quickly.  While the basic idea of perceptron based prefetch
filtering is applicable to any look-ahead prefetcher, we develop a
practical implementation of our proposed perceptron prefetch filter
using SPP as our baseline.  Here we describe the basic architecture of
SPP.

>>>>

Kim {\em et. al.} proposed Signature Path Prefetcher (SPP)~\cite{SPP},
a confidence-based lookahead prefetcher.  SPP creates a signature
associated with a page address by compressing the history of
accesses. By correlating the signature with future likely delta
patterns, SPP learns both simple and complicated memory access
patterns quickly.  While the basic idea of perceptron based prefetch
filtering is applicable to any lookahead prefetcher, we develop a
practical implementation of our proposed perceptron prefetch filter
using SPP as our baseline.  Here we describe the basic architecture of
SPP.

### 2018-12-07 intro.tex

\begin{figure}[t]
\includegraphics[width=\columnwidth]{Motivation}
\caption{Impact of aggressive prefetching on performance. X-axis
  represents increasingly aggressive prefetch lookahead depth. All the
  metrics have been normalized to the least aggressive
  configuration. As can be seen, total number of prefetches increase
  at a higher rate than the number of good prefetches, leading to IPC
  reduction.}
\label{Fig:Motivation}
\end{figure}

Coverage and accuracy are generally at odds with one another, and as
one metric improves, the other usually gets worse. For example, when
an application accesses a new region of memory for the first time, a
na\"ive prefetcher may predict that all data in that region will be
used by the application.  This will clearly result in 100\% coverage
for that region, but with possibly a very low accuracy.  In fact, so
much cache capacity and bandwidth may be wasted prefetching unused
data that performance can ultimately be harmed by this strategy. On
the other extreme, another prefetcher may be overly conservative and
never prefetch anything, wasting no capacity or bandwidth, and
achieving 0\% prefetch coverage.

Figure~\ref{Fig:Motivation} illustrates the above scenario.  Here we
consider a state-of-the-art. lookahead prefetcher -- SPP~\cite{SPP}.
Lookahead prefetchers such as SPP provide a mechanism to speculate an
arbitrary number of references ahead of the initial triggering miss.
In SPP, a throttling confidence threshold is then used to ensure that
the lookahead stops when confidence falls too low to ensure that
prefetches are accurate.  In the figure, we iteratively re-tuned 
this threshold and allowed the prefetcher to lookahead an of 
arbitrary depth from 7 to 15. The figure depicts the behavior 
of the {\tt 603.bwaves\_s} SPEC CPU 2017 benchmark. The IPC, the 
total number of prefetches suggested by the prefetcher (TOTAL\_PF) 
and the actual useful predictions (GOOD\_PF), all have been normalized 
to the least aggressive version of the prefetcher. As the look-ahead
depth increases, so do useful prefetches, and hence coverage. This
coverage, however, comes at a cost of total prefetches increasing at
an even higher rate. This leads to cache pollution and bandwidth
contention, as is clearly visible with the reduction in IPC.  
%Similar behavior was observed in 12 other SPEC CPU 2017 benchmarks.

>>>>

\begin{figure}[t]
\includegraphics[width=\columnwidth]{Motivation}
\caption{The impact of aggressive prefetching on performance for {\tt 603.bwaves\_s}. 
The number of useful prefetches increases with aggressiveness
slower than total prefetches, which wastes bandwidth and 
harms performance.}
\label{Fig:Motivation}
\end{figure}

Coverage and accuracy are generally at odds with one another, and as
one metric improves, the other usually gets worse. For example, when
an application accesses a new region of memory for the first time, a
na\"ive prefetcher may predict that all data in that region will be
used by the application.  This will clearly result in 100\% coverage
for that region, but with possibly a very low accuracy.  In fact, so
much cache capacity and bandwidth may be wasted prefetching unused
data that performance can ultimately be harmed by this strategy. At
the other extreme, another prefetcher may be overly conservative and
never prefetch anything, wasting no capacity or bandwidth, and
achieving 0\% prefetch coverage.

Figure~\ref{Fig:Motivation} illustrates the above scenario.  Here we
consider a state-of-the-art. lookahead prefetcher -- SPP~\cite{SPP}.
Lookahead prefetchers such as SPP provide a mechanism to speculate an
arbitrary number of references ahead of the initial triggering miss.
In SPP, a throttling confidence threshold is then used to ensure that
the lookahead stops when confidence falls too low to ensure that
prefetches are accurate.  In the figure, we iteratively re-tuned 
this threshold to allow the prefetcher to lookahead a fixed
depth from 7 to 15. The figure depicts the behavior 
of the {\tt 603.bwaves\_s} SPEC CPU 2017 benchmark. The IPC, the 
total number of prefetches issued by the prefetcher (TOTAL\_PF), 
and the actual useful predictions (GOOD\_PF), all have been normalized 
to lookahead depth 7. As the lookahead
depth increases, so do useful prefetches, and hence coverage. This
coverage, however, comes at the cost of total prefetches increasing at
an even higher rate. This leads to cache pollution and bandwidth
contention, and leads to a reduction in IPC.  
%Similar behavior was observed in 12 other SPEC CPU 2017 benchmarks.
